
%I need to fix the table size, add references to the error checking, add the number of non-tenrecs or golden mole skulls that I used, extra pictures available on request or put them onto figshare?

\chapter{Data appendix}
\label{appendix}

	The analyses in this thesis only use a subset of the total morphological data that I collected. This appendix describes how I collected the additional data. I include details on data from specimen labels, linear measurements of limbs and skulls and additional photographs of skulls, limbs and skins. These data represent significant resources which could be used for future research.

%-------------------------------------------------
\section{Specimen labels}

	I recorded all the data on the specimen labels including any handwritten or printed notes which had been added by other users of the collection. The label data included the museum specimen accession number, Genus, species, sex, collector's name, and the date and location of collection. Some of the labels attached to skins had additional information such as the body, tail, hind foot and ear lengths as well as the body mass of the live individual. 
%-------------------------------------
\section{Linear measurements}
\label{sect:measurements}

	My thesis focuses on geometric morphometric analyses of skull shape. However, I also collected extensive data based on linear measurements of mandibles, mandibles and limbs. 
	
	I used a 15 mm digital calipers (Mitutoyo Absolute digimatic calipers) for all of the measurements. I took five measurements of each mandible (table \ref{tab:mands.measurements}), 15 of the skulls (table \ref{tab:sk.measurements}) and 19 measurements of limbs (table \ref{tab:limb.measurements}). My choice of  measurements was based on three main criteria; 1) their relevance to biological and ecological traits such as diet specialisation and locomotory adaptation; 2) their usefulness for assessing the overall shape and size of the specimen; and 3) the ease with which they could be repeated both within and among specimens from different species. 

	I took each linear measurement three times, cycling through all of the measurements then repeating the cycle to avoid measuring the same variable twice in a row. Small measurements (<2 mm) are particularly prone to high error rates \citep{Cardini2008}. Therefore, I took five separate replicates of some of the measurements which were often less than 2 mm and consequently most prone to errors (marked with * in tables \ref{tab:sk.measurements} and \ref{tab:limb.measurements}). These included four of the skull measurements (PWa, IncisorH, IFD and IFcanal, table \ref{tab:mands.measurements}) and five of the limb measurements (FemD, TibD, HumD, UlnD, RadD, table \ref{tab:limb.measurements}). 
	Five replicates should give a more reliable median value because even if there are one or two outlying measurements there should be at least three replicates which are in close agreement \citep{Cooper2009}.

%---------------------------------------------
% Mandible  measurements
\begin{table}[!htbp]
	\caption[Mandible measurements]
			{Measurement abbreviations and descriptions for the mandibles, all taken from the labial (outer) side of the right jaw unless that side was broken or missing. All measurements were repeated three times.}
	\input{Methods/tables/mands.measurements}
	\label{tab:mands.measurements}
\end{table}

%NB: I need to fix the table lengths to go over more than one page

%I used the longtablepackage  to go over more than one page but I need to get \setcounter{LTchunksize}{10} working to change the number of rows that are involved

%Skull measurements
\begin{table}[!htbp]
	%\caption[Skull measurements]
			%{Measurement abbreviations and descriptions for the skulls. All measurements were repeated three times except for those marked with * which were measured five times.}% add to this caption
	\input{Methods/tables/skulls.measurements}
	%\label{tab:sk.measurements}
\end{table}





% Limb measurements
\begin{table}[!htbp]
	\caption[Limb measurements]
		{Measurement abbreviations and descriptions for the limbs. All measurements were repeated three times except for those marked with * which were measured five times.} % add to this caption
	\input{Methods/tables/limbs.measurements}
	\label{tab:limb.measurements}
\end{table}


%****************************
\newpage
\subsection{\normalfont{Error checking for linear measurements}}
	
	As mentioned above (section \ref{sect:measurements}), I took three replicate measurements of most of my variables and five replicates of other, smaller variables. 
	Some morphometric studies take replicate measurements of a trait and use the average value for further analyses. Rather than taking the mean of each of three (or five) measures, I used the median as it is less likely to be skewed by outliers and gives a more accurate representation of the true value of the trait.
	
	%Come back to here for references
	
	Before extracting the median values, I followed the protocol for assessing measurement error outlined by \citep{Cooper2009}. This method assesses whether there is a reasonable correlation among the replicate measurements of the same variable. The error checking criteria are based on two calculations; the coefficient of variation and the percentage spread.
	
	I calculated the coefficient of variation (standard deviation/mean*100) for each measurement. This value estimates the extent to which replicate measurements deviate from the mean. When the coefficient of variation was less than 5\%, I accepted the median value as an accurate measurement of the size of the structure. 
	If the coefficient of variation was greater than 5\%, indicative of a low agreement between replicate measurements, I measured the percentage spread of the data. For variables measured three times, I calculated percentage spread as [(minimum difference between neighbouring measurements)/ (range of measured values)*100].
	For variables that I measured five times, the differences between neighbouring values were calculated and labelled from smallest to largest as a, b, c, and d with the range of the measured values designated as e \citep{Cooper2009}. For these variables, I calculated percentage spread as [(a/e + b/e + c/e)*100]. 
	Small percentage spread values indicate close agreement between repeated measurements. When percentage spread approaches 50\% the data are evenly spread out and therefore there is no way of knowing whether the median value is an accurate measurement of the trait \citep{Cooper2009}. I chose to use to use 25\% as a cut off point for accepting the accuracy of measured traits.

	I used these error checking criteria to assess the accuracy of my repeated measurements of both skulls and limbs. 

%--------------------------------------
\section{Additional photographs}
	\subsection{\normalfont{Skulls}}
	Chapter \ref{chap:methods} describes my geometric morphometric analyses of tenrec and golden mole skulls. I used exactly the same photography set up and landmarks to analyse the photos of the non-Afrosoricida specimens (see table \ref{tab:species.measured} for reference). The TPS files containing landmark coordinates for this larger data set could be used  to measure morphological similarities among tenrecs and the species they convergently resemble. % Put in reference to discussion?
	%Find how many additional skulls/ species they include
		
	\subsection{\normalfont{Limbs}}
	Initially, I tried to take pictures of the limbs in similar orientations to the skulls (dorsal, ventral and lateral). However, there was considerable variation in how the limbs were preserved. For example, some limbs were still articulated while others had fragmented bones. It therefore proved impossible to place the limb bones in consistent orientations that would be comparable across species. Similarly, the small size of some limbs, combined with the frequently incomplete nature of postcranial museum collections, made landmark-based morphometric analyses of any limb pictures impractical. Therefore, I photographed the fore- and hind-limb bones in outer (the side facing away from the rest of the body) and inner (the side facing in towards the centre of the body) views for reference purposes only.	
	
	\subsection{\normalfont{Skins}}
	As I was limited by the maximum camera height available on the copy stands, most skins were too large to be photographed with the 100mm macro lens. Therefore, I used an EFS 18-55mm lens to take pictures of the skins. I photographed skins in the same three orientations as the skulls; dorsal (the upper surface of the animal), ventral (the belly side of the skin) and lateral (right flank of the animal with the skin held in position using bean bags). The dorsal and ventral views give very approximate estimates of the overall body shape of the animal. The lateral views are less biologically relevant since the taxidermic process is unlikely to produce specimens which represent the true body height of the animal. 
	There were four specimens from the Smithsonian Institute that had species labels which did not match between skulls and skins with the same specimen ID numbers. The four skulls were labeled as \textit{Hemicentetes semispinosus}. The corresponding skins were originally labeled as \textit{H. semispinosus} but this was crossed out and changed to \textit{H. nigriceps}. The re-labeled skins looked clearly different to the undisputed \textit{H. semispinosus} skins and also look more similar to other pictures of \textit{H. nigriceps}. Therefore, I made the assumption that the re-labeling of the skins as \textit{H. nigriceps} represents the true taxonomy and I treated the corresponding skulls as \textit{H. nigriceps}.
	
	%Images available on request or add them to figshare?

%--------------------------------------------

	

	
